---
title: "cm008 Exercises: Fix the Plots"
output: 
  html_document:
    keep_md: true
    theme: paper
---

In this worksheet, we'll be looking at some erroneous plots and fixing them. 

I think you might not have these two packages installed:

```
install.packages("ggridges")
install.packages("scales")
```

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(gapminder)
library(ggridges)
library(scales)
```

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```


## Exercise 1: Overlapping Points

After fixing the error, fix the overlapping problem in the following plot (attribution: ["R for data science"](https://r4ds.had.co.nz/data-visualisation.html)).

whenever youre working with ggplot, always +, geom jitter seperates points a bit, smooth makes line of best fit, linesar regression line. it is fitting it to originla points, not geom jitter. geom jitter doesnt actually change the actual data. geom smooth is using city of linear data the main data.



```{r, fig.width = 5, fig.height = 2}
ggplot(mpg, aes(cty, hwy)) +
  geom_point() +
  geom_smooth(method="lm")

ggplot(mpg, aes(cty, hwy)) +
  geom_jitter(alpha=0.5,size=1)+
  geom_smooth(method="lm")

ggplot(mpg, aes(cty, hwy)) +
  geom_point() +
  geom_jitter(alpha=0.5,size=1)+
  geom_smooth(method="lm")
```


## Exercise 2: Line for each Country

Fix this plot so that it shows life expectancy over time _for each country_. Notice that `ggplot2` ignores the grouping of a tibble! trying to show life expectancy , but can't 
grouping within ggplot is different than in dplyr, so it isnt recognizing it.
was geomline() but then all too dark, so alpha transparency of 0.2, could colour by country BUT NOT ENOUGH COORs. other ways to split it up. 
```{r, fig.width = 5, fig.height = 2}
gapminder %>% 
  group_by(country) %>% 
  ggplot(aes(year, lifeExp)) +
  geom_line()

gapminder %>% 
  #group_by(country) %>% 
  ggplot(aes(year, lifeExp,group=country)) +
  geom_line(alpha=0.2)

gapminder %>% 
  #group_by(country) %>% 
  ggplot(aes(year, lifeExp,group=country,color=country)) +
  geom_line(alpha=0.2)

gapminder %>% 
  #group_by(country) %>% 
  ggplot(aes(year, lifeExp,group=country,color=country=="Rwanda")) +
  geom_line(alpha=1/3)

#will give us a series of try and false. so if true, country is RWANDA, is blue.

gapminder %>% 
  #group_by(country) %>% 
  ggplot(aes(year, lifeExp,group=country,color=country=="Rwanda")) +
  geom_line(alpha=1/3)+
  scale_colour_discrete("LegendTitle", labels=c("Other","Rwanda"))
```


## Exercise 3: More gdpPercap vs lifeExp

### 3(a) Facets

- Change the x-axis text to be in "comma format" with `scales::comma_format()`.
- Separate each continent into sub-panels.
little c returns a vector
facet_wrap makes a panel for each of a categorical variable.
y axis on row 1 is same as y axis on row 2
if you want to zoom in, each thing have it's own axis , us e free. if you do scales="free+y" only y is variable. or with x, only x is 
```{r, fig.width = 5, fig.height = 2}
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(alpha = 0.2) +
  scale_x_log10(labels=scales::comma_format()) +
  facet_wrap(~continent)

ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(alpha = 0.2) +
  scale_x_log10(labels=scales::comma_format()) +
  facet_wrap(~continent, scales="free")
```

### 3(b) Bubble Plot

- Put the plots in one row, and free up the axes.
- Make a bubble plot by making the size of the points proportional to population. 
  - Try adding a `scale_size_area()` layer too (could also try `scale_radius()`).
- Use `shape=21` to distinguish between `fill` (interior) and `colour` (exterior). 
changenrow=1 added it to (~continent). added size=pop
missed why scale size area. dont need color because between continents.
but lets do it.

why adding aes in ggplot, and not do geom. in gplot, applies globally. if only ggpoint, only applies to that layer. 

```{r, fig.width = 8, fig.height = 2}
gapminder %>% 
  filter(continent != "Oceania") %>% 
  ggplot(aes(gdpPercap, lifeExp, size=pop)) +
  facet_wrap(~ continent, nrow=1) +
  geom_point(alpha = 0.1) +
  scale_x_log10(labels = scales::comma_format())


gapminder %>% 
  filter(continent != "Oceania") %>% 
  ggplot(aes(gdpPercap, lifeExp, size=pop, color=continent)) +
  facet_wrap(~ continent, nrow=1) +
  geom_point(alpha = 0.8) +
  scale_x_log10(labels = scales::comma_format())+
  scale_size_area()

gapminder %>% 
  filter(continent != "Oceania") %>% 
  ggplot(aes(gdpPercap, lifeExp, size=pop, color=continent)) +
  facet_wrap(~ continent, nrow=1) +
  geom_point(alpha = 0.8,shape=21) +
  scale_x_log10(labels = scales::comma_format())+
  scale_size_area()

#fill the points with color
gapminder %>% 
  filter(continent != "Oceania") %>% 
  ggplot(aes(gdpPercap, lifeExp, size=pop, fill=continent)) +
  facet_wrap(~ continent, nrow=1) +
  geom_point(alpha = 0.8,shape=21) +
  scale_x_log10(labels = scales::comma_format())+
  scale_size_area()
```

A list of shapes can be found [at the bottom of the `scale_shape` documentation](https://ggplot2.tidyverse.org/reference/scale_shape.html).

### 3(c) Size "not working"

Instead of alpha transparency, suppose you're wanting to fix the overplotting issue by plotting small points. Why is this not working? Fix it.

```{r, fig.width = 5, fig.height = 2}
ggplot(gapminder) +
  geom_point(aes(gdpPercap, lifeExp, size = 0.1)) +
  scale_x_log10(labels = scales::dollar_format())
```



## Exercise 4: Walking caribou

The following mock data set marks the (x,y) position of a caribou at four time points. 

- Fix the plot below so that it shows the path of the caribou. 
- Add an arrow with `arrow = arrow()`.
- Add the `time` label with `geom_text()`.

```{r, fig.width = 5, fig.height = 2}
tribble(
  ~time, ~x, ~y,
  1, 0.3, 0.3,
  2, 0.8, 0.7,
  3, 0.5, 0.9,
  4, 0.4, 0.5
) %>% 
  ggplot(aes(x, y)) + 
  geom_line()
```

## Exercise 5: Life expectancies in Africa

### 5(a) Unhiding the data

Fix the plot so that you can actually see the data points. Be sure to solve the problem of overlapping text, without rotating the text.

```{r, fig.width = 5, fig.height = 2}
gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot(aes(country, lifeExp)) + 
  geom_point() +
  geom_boxplot()
```

### 5(b) Ridgeplots

We're starting with the same plot as above, but instead of the points + boxplot, try a ridge plot instead using `ggridges::geom_density_ridges()`, and adjust the `bandwidth`.

```{r, fig.width = 5, fig.height = 2}
gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot(aes(country, lifeExp)) + 
  geom_point() +
  geom_boxplot()

gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot(aes(country, lifeExp)) + 
  ggridges::geom_density_ridges()

#upset because it was vertical, want to change the order of x and y because "missing aes"

gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot(aes(lifeExp,country)) + 
  ggridges::geom_density_ridges()
```

## Exercise 6: Bar plot madness

### 6(a) Colour and stacking madness

- Change the following plot so that it shows _proportion_ on the y-axis, not count.
- Change the x-axis so that it doesn't appear to be continuous.
- Put the bars for transmission side-by-side with their own colour.
- Capitalize the legend title.

```{r, fig.width = 5, fig.height = 2}
mtcars %>% 
  mutate(transmission = if_else(am == 0, "automatic", "manual")) %>% 
  ggplot(aes(cyl)) +
  geom_bar(aes(colour = transmission))
```

### 6(b) Bar heights already calculated

Here's the number of people having a certain hair colour from a sample of 592 people:

```{r}
(hair <- as_tibble(HairEyeColor) %>% 
  count(Hair, wt = n))
```

Fix the following bar plot so that it shows these counts.

```{r}
ggplot(hair, aes(Hair, n)) +
  geom_bar()
```

## Exercise 7: Tiling

Here's the number of people having a certain hair and eye colour from a sample of 592 people:

```{r}
(hair_eye <- as_tibble(HairEyeColor) %>% 
  count(Hair, Eye, wt = n))
```

Fix the following plot so that it shows a filled-in square for each combination. 

```{r}
ggplot(hair_eye, aes(Hair, Eye)) +
  geom_point(aes(colour = n))
```

By the way, `geom_count()` is like `geom_bar()`: it counts the number of overlapping points.


## Additional take-home practice

If you'd like some practice, give these exercises a try

__Exercise 1__: Make a plot of `year` (x) vs `lifeExp` (y), with points coloured by continent. Then, to that same plot, fit a straight regression line to each continent, without the error bars. If you can, try piping the data frame into the `ggplot()` function.

__Exercise 2__: Repeat Exercise 1, but switch the _regression line_ and _geom\_point_ layers. How is this plot different from that of Exercise 1?

__Exercise 3__: Omit the `geom_point()` layer from either of the above two plots (it doesn't matter which). Does the line still show up, even though the data aren't shown? Why or why not?

__Exercise 4__: Make a plot of `year` (x) vs `lifeExp` (y), facetted by continent. Then, fit a smoother through the data for each continent, without the error bars. Choose a span that you feel is appropriate.

__Exercise 5__: Plot the population over time (year) using lines, so that each country has its own line. Colour by `gdpPercap`. Add alpha transparency to your liking. 

__Exercise 6__: Add points to the plot in Exercise 5.