---
title: "cm009 Exercises: tidy data"
output: 
  html_document:
    keep_md: true
    theme: paper
---

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
lotr  <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/lotr_tidy.csv")
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

Tibble function, specify names

Create a name column. and then 
if you are using base r, 
can use trubble function to manually specify. 
list_col tibble. 
expand and explore grid-all combination of variables. all combinations of x =1-2, y=1-2. 
nesting if you dont want x and y lumped together. 
go through implicit NAs on your own. time series tibble where you go through missing date. same idea here. 

Untidy example 1

if tidy format, if different types, all you have to do is group by hair, and then summarize by eyes here. Example 2 , same thing. switched eye and hair color. there are wider. there are more columns as untidy, and there are less rows. untidy lots of columns, but fewer rows. need to reverse this, put it in different format. data that is too wide is not always issue. example data too long, and also untidy, because observations are split into two rows. going to need to group, and then do some spreading. each unique part of body part has a variable, so it should be in its own column, because it is a variable.

does red hair ever occur with blue eyes. 
can filter(hair == "red", eye=="blue")

there isnt such thing as wide format and long format. you can only prepare two formats, and say table one is longer than table two, or table one is wider than table two.

each variable is contained within a single column. what we want to do is take all of these variables and put them into one column
9.2., pivoting tools. pivot wider or pivot takes tibble and makes it either wider or longer.

reshaped package. reshape 2 package. actually pretty convenient to use gather and spread. terminology, gives you a lot more flexibility for widening or funcitonin. its not like the gather and spread functions are going away, just not actively___ not sure. 

what the fuck is gather and spread.
pivoting- multivariate pivoting.
eye color as columns, and we need to move this into a single column. need to take the titles, column names. these are actually values of an observation, so i need to take those into a single row. then i need to take whats underneath, and put those in a single row.

one column names, another column , calues. so names are now in "eye column" and values are in the n column.

pivoting-pivot longer to take the wide format and make it longer. takes a tibble and returns a tibble. besides the first argument that .... the tibble . we have to specify which columns are we working with . so we want all of the eye colour column. thats enough . optional and highly recommended are name to and values_to argument. the names are going to its own column. what do you want that column name to be . and then values too. the values that will be in ___ columns. taking the tidy data, making it longer. saying you want all of these columns green_eye, brown_eye,  and then specify that you want the names to be specified "eye".

I actually cant think of an example where you have two columns and make it wider. maybe if you had eye color and age. 
if there are 29 blue hair black eye peopl....we have list columns. tibbles with list columns. if you look at column blue. you have basically a list.

here, in specifying the columns, instead of - looked at the column to the left, and the column to right. still not entirely robust. tidy select- contain. contain eye. all of the columns i want to work with have the word iye in it

or all columns without hair, so all columns minus hair. that is lengthening. 

hair is going to be the identifyer. name scrub-names of the new column, which column is it coming vrom . and then values. so something needs to come underneath the values.

## Exercise 1: Univariate Pivoting

Consider the Lord of the Rings data:

```{r}
lotr
```

1. Would you say this data is in tidy format? relative

2. Widen the data so that we see the words spoken by each race, by putting race as its own column.

```{r}
(lotr_wide <- lotr %>% 
  pivot_wider(id_cols = c(-Race, -Words), 
              names_from = "Race", 
              values_from = "Words")
```

3. Re-lengthen the wide LOTR data from Question 2 above.

```{r}
lotr_wide %>% 
  pivot_longer(cols = c(-Film, -Gender), 
               names_to  = "Race", 
               values_to = "Words")
```

## Exercise 2: Multivariate Pivoting

9.4

Congratulations, you're getting married! In addition to the wedding, you've decided to hold two other events: a day-of brunch and a day-before round of golf.  You've made a guestlist of attendance so far, along with food preference for the food events (wedding and brunch).

```{r}
guest %>% 
  DT::datatable(rownames = FALSE)
```

1. Put "meal" and "attendance" as their own columns, with the events living in a new column.

```{r}
(guest_long <- guest %>% 
  pivot_longer(cols      = FILL_THIS_IN, 
               names_to  = FILL_THIS_IN,
               FILL_THIS_IN))
```

2. Use `tidyr::separate()` to split the name into two columns: "first" and "last". Then, re-unite them with `tidyr::unite()`.

```{r}
guest_long %>% 
  separate(FILL_THIS_IN, into = FILL_THIS_IN)
  # unite(col = "name", FILL_THIS_IN, sep = FILL_THIS_IN)
```

3. Which parties still have a "PENDING" status for all members and all events?

```{r}
FILL_THIS_IN %>% 
  group_by(party) %>% 
  summarize(all_pending = all(attendance == "PENDING"))
```

4. Which parties still have a "PENDING" status for all members for the wedding?

```{r}
FILL_THIS_IN %>% 
  group_by(party) %>% 
  summarize(pending_wedding = all(FILL_THIS_IN == "PENDING"))
```


5. Put the data back to the way it was.

```{r}
guest_long %>% 
  pivot_wider(id_cols     = c(party,name), 
              names_from  = event, 
              names_sep   = "_", 
              values_from = (meal,attendance))
```

6. You also have a list of emails for each party, in this worksheet under the variable `email`. Change this so that each person gets their own row. Use `tidyr::separate_rows()`
each row of guest column is just a bunch of names seperated by columns. each party has their own email.
seperate rows, comma space

```{r}
email %>% 
  separate_rows(guest, sep = ", ")
```


## Exercise 3: Making tibbles

1. Create a tibble that has the following columns:

- A `label` column with `"Sample A"` in its entries.
- 100 random observations drawn from the N(0,1) distribution in the column `x`
- `y` calculated as the `x` values + N(0,1) error. 
can use the tibble function
if vector length doesnt match, it is entry
rnorm function in R
tibble is backwayrds compatable. 
y, which is x that i just created.
create x, after data frame is created.

```{r}
n <- 100
FILL_THIS_IN(label = "Sample A",
             FILL_THIS_IN = rnorm(n),
             FILL_THIS_IN)
```


2. Generate a Gaussian sample of size 100 for each combination of the following means (`mu`) and standard deviations (`sd`).

```{r}
n <- 100
mu <- c(-5, 0, 5)
sd <- c(1, 3, 10)
FILL_THIS_IN(mu = mu, sd = sd) %>% 
  group_by_all() %>% 
  mutate(z = list(rnorm(n, mu, sd))) %>% 
  FILL_THIS_IN
```

3. Fix the `experiment` tibble below (originally defined in the documentation of the `tidyr::expand()` function) so that all three repeats are displayed for each person, and the measurements are kept. The code is given, but needs one adjustment. What is it? specify what is missing, so name, treatmen. but with complete function it is considering all possible combinations. we dont wat that. we want name and treatmen, and THEN rep

```{r}
experiment <- tibble(
  name = rep(c("Alex", "Robert", "Sam"), c(3, 2, 1)),
  trt  = rep(c("a", "b", "a"), c(3, 2, 1)),
  rep = c(1, 2, 3, 1, 2, 1),
  measurement_1 = runif(6),
  measurement_2 = runif(6)
)
experiment %>% expand(name, trt, rep)

experiment %>% complete(name, trt, rep)

experiment %>% complete(nesting(name, trt), rep)
```


