---
title: "cm010 Exercises: Tibble Joins"
output: 
  html_document:
    keep_md: true
    theme: paper
---

## Requirements

You will need Joey's `singer` R package for this exercise. And to install that, you'll need to install `devtools`. Running this code in your console should do the trick:

```
install.packages("devtools")
devtools::install_github("JoeyBernhardt/singer")
```

Load required packages:

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(singer)
knitr::opts_chunk$set(fig.width=4, fig.height=3, warning = FALSE, fig.align = "center")
```

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

## Exercise 1: `singer`


left join of x and y 
mutating join where we are adding columns to our base tibble, i thing right is x, and left is y? missed that

na is the , when there is nothing to combine, is no value.
result of join adds columns to your base nibble.
most common is left join.

anti join, gives you information that we don’t have in x.
y, just x dibble filtered out, so that person 3 is not there.

thats it for filtering joins. 
semi-join is basically just an inner join except you’re not keeping the columns alive.

if you have a data frame that has a tibble, you dont need to make into a tibble before using join function.

A=abc
B=abd
intersect, will take everything in common{ab}
union-will take all elements that appear in either of them. {a,b,c,d}
Set diff, A-B. starting with set A, elements are kept that dont appear in b {c}
set diff B-A woiudl be d left.

Bind_rows() and bind_cols

The package `singer` comes with two smallish data frames about songs. Let's take a look at them (after minor modifications by renaming and shuffling):

```{r}
(time <- as_tibble(songs) %>% 
   rename(song = title))
```

```{r}
(album <- as_tibble(locations) %>% 
   select(title, everything()) %>% 
   rename(album = release,
          song  = title))
```


1. We really care about the songs in `time`. But, which of those songs do we know its corresponding album?
for outer join we have left, righ, and (another ) join. these are mutating joins 9new row)
we want a filtering join. we like the time table. which of those do we have information about album and city
outer join.  filtering, semi join
each song will indentify bu observation
if two artists make same song name, would need to do by both song and artist

if we wanted to do same thing but display city and album info- mutate- inner join maybe

```{r}
time %>% 
  semi_join(album, by ="song", "artist_name")

time %>% inner_join by= c("song","artist_name") missed this

time %>% 
  semi_join(album, by ="song", "artist_name")
```

2. Go ahead and add the corresponding albums to the `time` tibble, being sure to preserve rows even if album info is not readily available.
left join preserve everything from base, and add if available (so doesnt throw away, does na)

```{r}
time %>% 
  left_join(album, by=c("song","artist_name"))
```

3. Which songs do we have "year", but not album info?
all that appear in time, but not in album.

```{r}
time %>% 
  semi_joinN(album, by = "song")
```

4. Which artists are in `time`, but not in `album`?
subtract out 
we are gonna see what is in time that is not in album. identify by artist name.

```{r}
time %>% 
  anti_join(album, by = "artist_name")
```


5. You've come across these two tibbles, and just wish all the info was available in one tibble. What would you do?
because you are not saying artist name is same, we will split from x y. doesnt know corduroy city and artist so confused. so add artist name
```{r}
time %>% 
  full_join(album, by = "song")

time %>% 
  full_join(album, by = c("song","artist_name"))
```


## Exercise 2: LOTR

Load in the three Lord of the Rings tibbles that we saw last time:

```{r}
fell <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Fellowship_Of_The_Ring.csv")
ttow <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Two_Towers.csv")
retk <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Return_Of_The_King.csv")
```

1. Combine these into a single tibble.

```{r}
bind_rows(fell, ttow, retk)
```

2. Which races are present in "The Fellowship of the Ring" (`fell`), but not in any of the other ones?
by is the column that indicates the observation
get 0 rows again. even if first antijoin get rids of everything what isf you get more data. code is not so robust then.
```{r}
fell %>% 
  anti_join(ttow, by = "Race") %>% 
  anti_join(retk, by = "Race")
```

## Exercise 3: Set Operations

Let's use three set functions: `intersect`, `union` and `setdiff`. We'll work with two toy tibbles named `y` and `z`, similar to Data Wrangling Cheatsheet

```{r}
(y <-  tibble(x1 = LETTERS[1:3], x2 = 1:3))
```

```{r}
(z <- tibble(x1 = c("B", "C", "D"), x2 = 2:4))
```

1. Rows that appear in both `y` and `z`
trying inner join as well, it also works.
might have written innner join wrong

```{r}
intersect(y, z)
inner_join(y,z)
```

2. You collected the data in `y` on Day 1, and `z` in Day 2. Make a data set to reflect that.
he would prefer bind rows. two tibbles, dont want option of removing rows, if sasme data on day 1 and day 2, and then it is not removed

```{r}
union(
  mutate(y, day = "Day 1"),
  mutate(z, day = "Day 2")
)

bind_rows(
  mutate(y, day = "Day 1"),
  mutate(z, day = "Day 2")
)
```

3. The rows contained in `z` are bad! Remove those rows from `y`.
antijoin also works, because same columns, if same columns in z and y we would have to use
```{r}
setdiff(y, z)
anti_join(y,z)
```
